# Modelli di simulazione delle colture

Un modello, nella sua accezione più ampia, è una rappresentazione semplificata di un sistema ben delimitato nel mondo reale. In agricoltura, ad esempio, una coltura con i suoi organi e processi rappresenta un sistema complesso. Negli ultimi anni, i modelli di simulazione della crescita delle colture hanno acquisito un ruolo cruciale negli studi e nelle applicazioni agrometeorologiche, consentendo di analizzare il comportamento delle colture e la loro risposta all'ambiente circostante.

Una delle caratteristiche distintive di questi modelli è la loro capacità di fornire risultati validi anche in contesti agronomici, colturali e gestionali diversi da quelli in cui sono stati originariamente sviluppati. Questa versatilità ha elevato la modellistica a strumento complementare ideale alla sperimentazione tradizionale.

In generale, la costruzione di un modello consiste nell'identificazione di un insieme di equazioni matematiche che riproducano nel modo più accurato possibile il comportamento del sistema in esame.

## Chapters and sub-chapters

There are two steps to cross-reference any heading:

1.  Label the heading: `# Hello world {#nice-label}`.
    -   Leave the label off if you like the automated heading generated based on your heading title: for example, `# Hello world` = `# Hello world {#hello-world}`.
    -   To label an un-numbered heading, use: `# Hello world {-#nice-label}` or `{# Hello world .unnumbered}`.
2.  Next, reference the labeled heading anywhere in the text using `\@ref(nice-label)`; for example, please see Chapter \@ref(cross).
    -   If you prefer text as the link instead of a numbered reference use: [any text you want can go here](#cross).

## Captioned figures and tables

Figures and tables *with captions* can also be cross-referenced from elsewhere in your book using `\@ref(fig:chunk-label)` and `\@ref(tab:chunk-label)`, respectively.

See Figure \@ref(fig:nice-fig).

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center', fig.alt='Plot with connected points showing that vapor pressure of mercury increases exponentially as temperature increases.'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Don't miss Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(pressure, 10), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```

```{r}
library(tidyverse)
# Function to simulate soil water balance
soil_water_balance <- function(soil_params, rain, ET0, irrigation) {
  
  # Initialize soil water
  available_water <- soil_params$InitialAmountOfAvailableWater
  results <- data.frame(Day = 1:length(rain))
  
  # Loop through days
  for (day in 1:length(rain)) {
    
    # Calculate transpiration
    transpiration <- min(0.096 * available_water, ET0[day]) 
    
    # Calculate surface runoff (SCS method)
    pot_max_ret <- (25400 / soil_params$RunoffCurveNumber) - 254
    init_abs <- 0.2 * pot_max_ret
    surface_runoff <- ifelse(rain[day] <= init_abs, 0,
                             ((rain[day] - init_abs)^2) / (rain[day] - init_abs + pot_max_ret))
    
    # Calculate deep drainage
    avail_water_before_drain <- available_water + rain[day] + irrigation[day] - transpiration - surface_runoff
    deep_drainage <- ifelse(soil_params$WaterHoldingCapacity >= (avail_water_before_drain / soil_params$RootZoneDepth), 
                            0, 
                            soil_params$DrainageCoeff * soil_params$RootZoneDepth * 
                              ((avail_water_before_drain / soil_params$RootZoneDepth) - soil_params$WaterHoldingCapacity))
    
    # Update available water
    available_water <- available_water + rain[day] + irrigation[day] - transpiration - surface_runoff - deep_drainage
    
    # Store results
    results$Rain[day] <- rain[day]
    results$Deep_drain[day] <- deep_drainage
    results$Run_off[day] <- surface_runoff
    results$ET0[day] <- ET0[day]
    results$Transpi[day] <- transpiration
    results$Avlbl_water[day] <- available_water
  }
  
  return(results)
}

# Example usage (replace with your actual data)
soil_params <- list(
  InitialAmountOfAvailableWater = 100,  # mm
  RunoffCurveNumber = 85,
  RootZoneDepth = 1,                     # m
  DrainageCoeff = 0.1,
  WaterHoldingCapacity = 0.2            # mm/mm
)

# Generate weather data for a year (365 days)
set.seed(42)  # For reproducibility
days <- 1:365

# Rain: Simulate a seasonal pattern (higher in spring/fall)
rain <- 20 * abs(sin(2 * pi * days / 365)) + rnorm(365, 0, 5)  # Seasonal + random noise

# ET0: Simulate a seasonal pattern (higher in summer)
ET0 <- 5 + 3 * abs(sin(2 * pi * (days - 90) / 365)) 

# Irrigation: Two scenarios
irrigation_scenario1 <- rep(0, 365)  # No irrigation
irrigation_scenario2 <- ifelse(days >= 150 & days <= 200, 5, 0)  # 5 mm/day from June 1st to Sept 1st

# Run the simulations for both scenarios
results_scenario1 <- soil_water_balance(soil_params, rain, ET0, irrigation_scenario1)
results_scenario2 <- soil_water_balance(soil_params, rain, ET0, irrigation_scenario2)
```

```{r}
# Reshape data for plotting
library(tidyr)
results_long_scenario1 <- results_scenario1 %>%
  pivot_longer(-Day, names_to = "Variable", values_to = "Value") %>%
  mutate(Scenario = "No Irrigation")
results_long_scenario2 <- results_scenario2 %>%
  pivot_longer(-Day, names_to = "Variable", values_to = "Value") %>%
  mutate(Scenario = "Irrigation")
results_long <- rbind(results_long_scenario1, results_long_scenario2)

```

```{r}
# Plot the results using ggplot2
library(ggplot2)

p <- ggplot(results_long, aes(x = Day, y = Value, color = Scenario)) +
  geom_line() +
  facet_wrap(Variable~., scales = "free_y") +  # Separate plots for each scenario
  labs(title = "Soil Water Balance Simulation", x = "Day of Year", y = "Water (mm)") +
  scale_color_manual(values = c("Irrigation" = "blue", "No Irrigation" = "orange")) +
  # scale_color_manual(values = c("Rain" = "blue", "Deep_drain" = "orange", 
    #                            "Run_off" = "green", "ET0" = "red", 
    #                           "Transpi" = "purple", "Avlbl_water" = "brown")) +
  theme_minimal()

print(p)
```
